---
title: "Statistics. Exam-demo"
date: "17 March 2016"
output: pdf_document
---

<!--
## Problem 1

What is a variance of random variable? Please, give a formal definition and an intuitive explanation. 
-->

## Problem 1

The distribution of random variable $X$ is given by the table:

value of $X$ | 1 | 4 | -3
------------- | ------|---|---
probability  | $0.2$ | $0.3$ | ?

1. What is masked by "?" in the table?
2. Find $E(X)$, $E(X^2)$, $E(X^3)$
3. Find $Var(X)$ and $\sigma_X$

## Problem 2

The distribution of random variables $X$ and $Y$ is given by the table:

joint probabilities | $X=0$ | $X=-1$ | $X=2$
------------- | -----|----|----
$Y=0$ | $0.2$ | $0.1$ | $0.1$
$Y=2$ | $0.2$ | $0.3$ | $0.1$

1. Find $Var(X)$, $Var(Y)$
2. Find $Cov(X, Y)$
3. Find $Corr(X, Y)$

## Problem 3

The density function of a random variable $X$ is given by

\[
f(x) = \begin{cases}
\frac{2}{3}x, x \in [0;1.5] \\
4-2x, x \in [1.5;2] \\
0, \text{otherwise}
\end{cases}
\]

1. Draw the density function
2. Find $P(X>1)$
3. Find the theoretical median of $X$
4. What is bigger $E(X)$ or 1?

## Problem 4

Random variables $X$ and $Y$ are independent, $X\sim N(5; 10)$, $Y\sim N(-3;5)$.

1. Find $E(X-2Y)$, $Var(X-2Y)$
2. Find $Corr(X-2Y, X+3Y)$

## Problem 5

Consider a small random sample of 4 observations, $X_1 = 2$, $X_2=-3$, $X_3=4$, $X_4=1$.

Calculate sample mean, sample variance and sample median.

## Problem 6

Researcher Eleonora has sampled 900 small firms. The sample mean revenue was equal to 3000 dollars per week with sample standard deviation of 1000 dollars per week.

1. At 5% significance level test the hypothesis that the expected revenue is equal to 3100.
2. Build a 95% confidence interval for expected revenue.

## Problem 7

Researcher Veniamin wonders whether the type of pies he eats influences the number of solved problems. Here $y_i$ is the number of solved problems during the evening, $x_i$ --- number of apple-pies eaten, $z_i$ --- number of meat-pies eaten:

\[
y_i = \beta_1 + \beta_2 x_i + \beta_3 z_i + u_i
\]

```{r, include = FALSE}
library(pander)
set.seed(42)
x <- sample(x = 2:9, size = 100, replace = TRUE)
z <- sample(x = 1:6, size = 100, replace = TRUE)
eps <- sample(x = c(-2,-1,0,1,2), size = 100, replace = TRUE)
y <- 5 + 3 * x + 2 * z + eps
model <- lm(y~x+z)
```

```{r, echo = FALSE, results='markup'}
pander(model)
```


1. Interpret the coefficients before $x_i$
2. Test the significance of every coefficient at $\alpha = 0.1$ significance level
3. Calculate 95% confidence interval for coefficient $\beta_2$
4. Forecast $y$ for $x=3$ and $z=7$

## Problem 8

Consider two regression models:

Model A:
\[
\hat y_i = 90 + 2.8 x_i 
\]

Model B:
\[
\hat y_i = -80 + 3.1 x_i - 2.6 z_i + 2.9 w_i
\]

The F-test for these two models gave P-value of 0.0002016.

1. Describe the F-test for nested model comparison: What is the $H_0$ for the test? The alternative hypothesis $H_a$? 
2. Which model do you prefer in this particular case and why?

## Problem 9

Researcher Veniamin would like to predict whether it will rain tomorrow ($y_i$) given the number of times his cat has meowed today ($x_i$). Veniamin has estimated the logit model:

\[
P(y_i = 1) = F(\beta_1  + \beta_2 x_i)
\]

Coefficient  | Estimate | Std. error
------------- | -------|------
$\hat\beta_1$  | 5 | 2.8
$\hat\beta_2$ | -2 | 0.5


1. What is the probability of rain if the cat has meowed 3 times?
2. Is the coefficient $\hat\beta_2$ significantly different from zero at 5% significance level?


## Problem 10

You have 4 observations on weight, $x_i$ (kg), and ice-cream preference ($y_i$):

$x_i$ | 60 | 80 | 90 | 70
----- | ---|---|---|---
$y_i$ | chocolate | vanilla | chocolate | chocolate


1. What is the Gini index?
2. Build a classification tree for $y$ using $x$ as predictor
